---
title: "Untitled"
author: "Bruce Nguyen"
date: "4/9/2021"
output: html_document
---

```{r}
knitr::knit_engines$set(python = reticulate::eng_python)
```

## Preparation

Here are the libraries that we need.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(GGally)
library(CCA)
library(tidyverse)
library(boot)
library(car)
library(ggrepel)
library(latex2exp)
library(MVN)
```

Read the data.

``` {r}
df <- read.csv("econ-hdi.csv")
df[1:10,]
```

```{r}
summary(df)
```



```{r}
econ <- df[,3:14]
hdi <- df[,15:18]
```


Measure multicollinearity. This is because it can have an effect on results' interpretation.

```{r}
lin_reg <- lm(Gross.national.income..GNI..per.capita ~., data = df[,3:18])

car::vif(lin_reg)
```

## Testing for independence between the 2 sets

* Note: The testing method used here is different from the one that is used in the report.

```{r}
ind_test <- function(d, i){
	d2 <- d[i,]
	econ <- d2[,3:14]
	hdi <- d2[,15:18]
	p <- ncol(econ)
	q <- ncol(hdi)
	m <- min(p,q)
	n <- nrow(d2)
	cc_sample <- cc(econ,hdi)
	test_stat <- -(n-0.5*(p+q+3))*sum(log(1-cc_sample$cor^2))
	return(test_stat)
}

partial_ind_test <- function(d,i) {
	d2 <- d[i,]
	econ <- d2[,3:14]
	hdi <- d2[,15:18]
	p <- ncol(econ)
	q <- ncol(hdi)
	m <- min(p,q)
	n <- nrow(d2)
	cc_sample <- cc(econ,hdi)
	test_stat <- -(n-0.5*(p+q+3))*sum(log(1-cc_sample$cor^2))
	return(test_stat)
}

results <- boot(data=df, statistic=ind_test, R=1000)
quant <- quantile(results$t,c(0.025,0.975))
(results$t0 > quant[1]) & (results$t0 < quant[2])
```

## Canonical Correlation Analysis

For detailed analysis, please look at the report.

```{r}
cc1 <- cc(econ, hdi)
cc1.cls_ind <- c("X", "Y")[c(rep(1,ncol(econ)),rep(2,ncol(hdi)))]
cc1.coefs <- as.data.frame(rbind(cc1$xcoef,cc1$ycoef))
cc1.coefs$Set <- cc1.cls_ind
cc1.coefs
```

```{r}
plot(cc1$cor, type="b", ylab="Value", xlab="Index of Canonical Correlations")
```

```{r}
cc1.ycoefs <- as.data.frame(cc1$ycoef)

gg <- ggplot(cc1.ycoefs, aes(x=V1, y=V2)) + 
  geom_point(size=3) +
  labs(color='Variables Set') + 
  labs(title="Canonical Correlation Coefficients for set Y",
       y=TeX(r'($\beta_2$)'),
       x=TeX(r'($\beta_1$)')) + 
  geom_text_repel(aes(label = rownames(cc1.ycoefs)),size=3,alpha=0.5,max.overlaps = 12)

plot(gg)
ggsave(device='png', filename = "coefY.png",dpi=700)
```


```{r}
cc1.xcoefs <- as.data.frame(cc1$xcoef)

gg <- ggplot(cc1.xcoefs, aes(x=V1, y=V2)) + 
  geom_point(size=3) +
  labs(color='Variables Set') + 
  labs(title="Canonical Correlation Coefficients for set X",
       y=TeX(r'($\alpha_2$)'),
       x=TeX(r'($\alpha_1$)')) + 
  geom_text_repel(aes(label = rownames(cc1.xcoefs)),size=3,alpha=0.5,max.overlaps = 12)

plot(gg)
ggsave(device='png', filename = "coefX.png",dpi=700)

```


```{r}
cc1 <- cc(econ, hdi)
cc1.cls_ind <- c("X", "Y")[c(rep(1,ncol(econ)),rep(2,ncol(hdi)))]
cc1.uv <- as.data.frame(rbind(cc1$scores$corr.X.xscores,cc1$scores$corr.Y.yscores))
cc1.uv$Set <- cc1.cls_ind
cc1.uv
# dot_col <- c("blue","red")[cc1.cls_ind]
# legends <- c("Economic Freedom", "HDI Index")[cc1.cls_ind]
# # plot the coefficients of each X
# plot(cc1.uv[,1],cc1.uv[,2],pch = 16, col = dot_col, cex = 1.25)
# text(cc1.uv[,1],cc1.uv[,2], labels=colnames(econ), pos=2, cex=0.7)

gg <- ggplot(cc1.uv, aes(x=V1, y=V2)) + 
  geom_point(aes(col=Set), size=3) + 
  labs(color='Variables Set') + 
  labs(title="Canonical Correlation Variant",
       y=TeX(r'($u_2$ (for X) and $v_2$ (for Y))') , x=TeX(r'($u_1$ (for X) and $v_1$ (for Y))')) + 
  geom_text_repel(aes(label = rownames(cc1.uv)),size=3,alpha=0.5,max.overlaps = Inf)

plot(gg)
ggsave(device='png', filename = "uv.png",dpi=700)
```



```{r}

cc1.scores <- as.data.frame(cbind(cc1$scores$xscores[,1],cc1$scores$yscores[,1]))

gg <- ggplot(cc1.scores, aes(x=V1, y=V2)) + 
  geom_point(aes(col=df[,ncol(df)]), size=2) + 
  labs(color='Country classification') + 
  labs(title="Scores (Opposite Sign)",
       y="Economic Freedom" , x="Human Development") + 
  geom_text_repel(aes(label = df[,2]),size=3,force=0.1,alpha=0.5,max.overlaps = 5)
  # geom_text(aes(label=df[,2]),hjust=1, vjust=1, size=3, alpha=0.5)

plot(gg)
ggsave(device='png', filename = "scores.png",dpi=700)
```


